---
title: "Elevated plasma lipopolysaccharide concentrations in pulmonary tuberculosis associate with immunometabolic outcomes"
author: "Zaynab Mousavian"
date: "September 28, 2026"
output:
  html_document: default
  pdf_document: default
subtitle: LPS data Analysis
---

```{r setup, include=FALSE}
# Set global options for all chunks
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)
```

# Load packages 

```{r}
library(gt)
library(readxl)
library(dplyr)
library(Hmisc)
library(reshape2)
library(circlize)
library(limma)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(ggpubr)
library(lme4)
library(broom.mixed)
library(purrr)
library(lmerTest)
library(ggplot2)
library(dplyr)
library(ggsignif)
library(tidyr)
library(dplyr)
library(circlize)
library(ComplexHeatmap)
library(corrplot)
library(igraph)
library(ggraph)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(circlize)
library(ComplexHeatmap)
library(grid)
library(ggbreak)
library(dplyr)
library(purrr)
library(broom)
library(gt)
library(flextable)
library(officer)
library(tibble)
```

# Functions {.tabset}
## Function for scatterplot of marker vs. cytokine

```{r}
plot_cyto_marker <- function(data, method, z, x_title="Plasma cytokine concentration",y_title="LPS (EU/ml)",n=7) {
  
  ggplot(data,
         aes(x = Cytokine, y = Marker, color = Group)) +
    
    geom_point(color="darkred") +
    
    
    facet_wrap(~ Pair, scales = "free", ncol = n) +
    stat_cor(
      data = subset(data),
 aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
 method = method,
      label.x.npc = 0.05,
      label.y.npc = 1-z,
      size = 4,
      color = "darkred"
    )+
    
    stat_cor(
      data = subset(data, Group == "Stress Hyperglycemia"),
   aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
   method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.94-z,
      size = 4,
      color = "#3abbcc"
    ) +
    
    stat_cor(
      data = subset(data, Group == "Euglycemia"),
    aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
    method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.88-z,
      size = 4,
      color = "#ca6332"
    )  +
    
    stat_cor(
      data = subset(data, Group == "Control"),
   aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
   method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.88-z,
      size = 3,
      color = "black"
    ) +
    
    labs(color = "", x = bquote(log[2](.(x_title))), y = bquote(log[10](.(y_title)))) +
    
    theme_classic()+
    theme(legend.position = "bottom")
}
```



## Function for scatterplot of marker vs. cytokine with time

```{r}
plot_cyto_marker_with_time <- function(data, method, z, x_title="Plasma cytokine concentration",y_title="LPS (EU/ml)",n=4) {
  

  ggplot(data,
         aes(x = Cytokine, y = Marker, color = Group)) +
    
    geom_point(color = "darkred") +
    
    
    facet_wrap(
      ~ Pair + time + time2,
      scales = "free",
      ncol = n,
      labeller = labeller(
        time = function(x)
          paste0("LPS: ", x),
        time2 = function(x)
          paste0("Cytokine: ", x)
      )
    )+
    stat_cor(data = subset(data),
             aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
 method = method,
      label.x.npc = 0.05,
      label.y.npc = 1-z,
      size = 6,
      color = "darkred"
    )+
    
    stat_cor(
      data = subset(data, Group == "Stress Hyperglycemia"),
   aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
   method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.94-z,
      size = 6,
      color = "#3abbcc"
    ) +
    
    stat_cor(
      data = subset(data, Group == "Euglycemia"),
    aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
    method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.88-z,
      size = 6,
      color = "#ca6332"
    )  +
    
    stat_cor(
      data = subset(data, Group == "Control"),
   aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
   method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.88-z,
      size = 6,
      color = "black"
    ) +
    
    labs(color = "", x = bquote(log[2](.(x_title))), y = bquote(log[10](.(y_title)))) +
    
    theme_classic() +
    theme(
      axis.text = element_text(size = 18),
      axis.title = element_text(size = 23),
      strip.text = element_text(face = "bold", size = 14),
      legend.position = "none")
}
```


## Function for scatterplot of LPS vs. marker without facet
```{r}
plot_LPS_marker_withoutfacet <- function(data, m, t, method) {
  
  data  <- data%>%filter(marker==m,time==t)
  
  
  ggplot(data, aes(x = value, y = LPS, color = Group)) +
    
    geom_point(color = "darkred", size = 3) +
    
    geom_smooth(
      method = "lm",
      se = FALSE,
      color = "darkred",
      linewidth = 1
    ) +
    
    stat_cor(
      data = subset(data),
      aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
      method = method,
      label.x.npc = 0.05,
      label.y.npc = 1,
      size = 6,
      color = "darkred"
    ) +
    
    stat_cor(
      data = subset(data, Group == "Stress Hyperglycemia"),
      aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
      method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.96,
      size = 6,
      color = "#3abbcc"
    ) +
    
    stat_cor(
      data = subset(data, Group == "Euglycemia"),
      aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
      method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.92,
      size = 6,
      color = "#ca6332"
    ) +
    
    scale_color_manual(
      values = c(
        "Stress Hyperglycemia" = "darkred",
        "Euglycemia" = "darkred"
      ),
      labels = c(
        "Stress Hyperglycemia" = "TB-stress hyperglycemia",
        "Euglycemia" = "TB-euglycemia"
      )
    ) +
    
    labs(
      color = "",
      y = bquote(log[10]("LPS (EU/ml)") ~ "at" ~ .(t)),
      x = paste0(m," at ",t)
    ) +
    
    theme_classic()+
    theme(axis.text = element_text(size=23),
         axis.title = element_text(size=23),
          legend.position = "none")
}
```

## Function for scatterplot of delta LPS vs. delta marker
```{r}
plot_deltaLPS_deltamarker_withoutfacet <- function(data, m, method) {
  
   data  <- data%>%filter(marker==m)
  ggplot(data,
         aes(x = deltaLPS, y = value, color = Group)) +
    
    geom_point(color="darkred",size=3)  +
    
 
    stat_cor(
      data = subset(data),
      aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
      method = method,
      label.x.npc = 0.05,
      label.y.npc = 1,
      size = 6,
      color = "darkred"
    ) +
    
    stat_cor(
      data = subset(data, Group == "Stress Hyperglycemia"),
      aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
      method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.96,
      size = 6,
      color = "#3abbcc"
    ) +
    
    stat_cor(
      data = subset(data, Group == "Euglycemia"),
      aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
      method = method,
      label.x.npc = 0.05,
      label.y.npc = 0.92,
      size = 6,
      color = "#ca6332"
    ) +
    
    scale_color_manual(
      values = c(
        "Stress Hyperglycemia" = "darkred",
        "Euglycemia" = "darkred"
      ),
      labels = c(
        "Stress Hyperglycemia" = "TB-stress hyperglycemia",
        "Euglycemia" = "TB-euglycemia"
      )
    ) +
    
    labs(
      color = "",
      x = expression(log[10]("LPS at Month 6 / LPS at baseline")),
      y = paste0("Relative change in ",m," from baseline to Month 6")
    ) +
    
    theme_classic()+
    theme(axis.text = element_text(size=23),
         axis.title = element_text(size=23),
          legend.position = "none")
}
```



## Fucntion for differential correlations between groups
```{r}
diff_correlation_test <- function(corr1, corr2, n1, n2) {
  
  results <- corr1 %>%
    rename(r1 = value, p1=p) %>%
    inner_join(
      corr2 %>% rename(r2 = value, p2=p),
      by = c("from", "to")
    ) %>%
    mutate(
      z1 = 0.5 * log((1 + r1) / (1 - r1)),
      z2 = 0.5 * log((1 + r2) / (1 - r2)),
    
      z_diff = (z1 - z2) / sqrt(1/(n1 - 3) + 1/(n2 - 3)),
      
      p_value = 2 * pnorm(-abs(z_diff)),
      
      FDR = p.adjust(p_value, method = "BH"),
      
      direction = case_when(
        r1 > r2 ~ "Higher in TB-stress hyperglycemia",
        r1 < r2 ~ "Higher in TB-euglycemia",
        TRUE ~ "Equal"
      )
    ) %>%
    arrange(FDR)
  
  return(results)
}
```


## Function for visaulizing heatmap plot 

```{r}
heatmapPlot_func <- function(ClinicalData, CytokineData,
                             name_col = "Cytokine",
                             id_col = "SAMPLE.ID",
                             group_col = "group",
                             lps_col = "LPS",
                             logFC_thr = 0,
                             qval_thr = 0.05,
                             up_color = "red",
                             mid_color = "white",
                             down_color = "blue",
                             annotation_colors = c(
                               "Euglycemia" = "#ca6332",
                               "Stress Hyperglycemia" = "#3abbcc"
                             ),
                             row_fontsize = 20,
                             col_fontsize = 20) {
  
  
  #----------------------------------
  # Sort samples by group
  #----------------------------------
  
  ClinicalData_sorted <- ClinicalData %>%
    arrange(.data[[group_col]])
  
  sample_sorted <- ClinicalData_sorted %>%
    pull(.data[[id_col]])
  
  
  #----------------------------------
  # Heatmap data
  #----------------------------------
  
  input_data <- CytokineData[, sample_sorted, drop = FALSE]
  
  input_data[is.na(input_data)] <- 0
  
  input_data_scaled <- t(apply(input_data, 1, function(x) {
    s <- scale(x)
    if (all(is.na(s))) rep(0, length(x)) else s
  }))
  
  colnames(input_data_scaled) <- colnames(input_data)
  
  
  #----------------------------------
  # Heatmap colors
  #----------------------------------
  
  color_fun <- colorRamp2(
    c(4, 0, -4),
    c(up_color, mid_color, down_color)
  )
  
  #----------------------------------
  # Annotations
  #----------------------------------
  
  annotation_group <- ClinicalData_sorted[[group_col]]
  annotation_lps   <- ClinicalData_sorted[[lps_col]]
  
  # Color functions
  lps_color_fun <- colorRamp2(
    c(
      min(annotation_lps, na.rm = TRUE),
      median(annotation_lps, na.rm = TRUE),
      max(annotation_lps, na.rm = TRUE)
    ),
    c("blue", "white", "red")
  )
  
  annotation_colors_list <- list(
    Group = annotation_colors,
    LPS = lps_color_fun
  )
  
  column_annotation <- HeatmapAnnotation(
    
    Group = annotation_group,
    LPS = annotation_lps,
    
    col = annotation_colors_list,
    
    annotation_name_gp = gpar(
      col = "black",
      fontsize = 20,
      fontface = "bold"
    ),
    
    show_legend = TRUE,
    
    annotation_legend_param = list(
      
      Group = list(
        title = "Group",
        title_gp = gpar(
          fontsize = 26,
          fontface = "bold"
        ),
        labels_gp = gpar(fontsize = 20)
      ),
      
      LPS = list(
        title = "LPS",
        title_gp = gpar(
          fontsize = 26,
          fontface = "bold"
        ),
        labels_gp = gpar(fontsize = 20)
      )
    )
  )
  
  
  #----------------------------------
  # Heatmap
  #----------------------------------
  
  plot <- Heatmap(
    input_data_scaled,
    name = "Z-score",
    
    row_names_side = "left",
    row_names_gp = gpar(
      fontsize = row_fontsize,
      col = "black",
      fontface = "bold"
    ),
    
    row_gap = unit(1.5, "mm"),
    
    column_names_gp = gpar(
      fontsize = col_fontsize
    ),
    
    cluster_columns = FALSE,
    column_gap = unit(2, "mm"),
    
    col = color_fun,
    
    cluster_rows = TRUE,
    show_row_dend = TRUE,
    
    show_column_names = FALSE,
    
    row_title = NULL,
    
    top_annotation = column_annotation,
    
    show_heatmap_legend = TRUE,
    
    heatmap_legend_param = list(
      title = "Z-score",
      title_gp = gpar(
        fontsize = 26,
        fontface = "bold"
      ),
      labels_gp = gpar(fontsize = 16)
    )
  )
  
  
  return(plot)
}
```

# Read data

## Read LPS data

```{r}
LPS_data <- read.csv("../Data/LPS_gut_markers.txt",sep = "\t")
LPS_data[which(LPS_data$LPS<0.0001),"LPS"] <- 0.006
LPS_data$SAMPLE.ID <- gsub("_","-",LPS_data$SAMPLE.ID)
# Exclude MC-0029 as Dzigbordi suggested
LPS_data <- LPS_data%>%filter(!is.na(LPS),SAMPLE.ID!="MC-0029")
table(LPS_data$TIME,LPS_data$Cohort)
```

## Add glycemic group to LPS data
```{r}
clinical_data <- read_excel("../Data/MIDSTData_AllPartic_GlycStat_11Jun2026.xlsx")
# clinical_data <- read_excel("../Data/MIDSTData_AllPartic_GlycStat_24Jul2026.xlsx") identical with this file, no update for LPS data
clinical_data <- clinical_data%>%rename(glycem_stat = glycem_stat_v7)
LPS_data$group <- clinical_data[match(LPS_data$SAMPLE.ID,clinical_data$record_id),"glycem_stat"]%>%unlist()
LPS_data[which(LPS_data$Cohort=="ADIEU"),"group"] <- "Control"
table(LPS_data%>%select(group,TIME))
```

## Add clinical outcomes to LPS data

```{r}
clinical_data_full <- read_excel("../Data/MIDST_ZMR21.xlsx")
clinical_data_control <- read_excel("../Data/NCTLDADIEUAdipokines-WassimMatheusData_DATA_LABELS_2024-11-11_0928.xlsx")
clinical_data_control$`Record ID` <- gsub("--1","",clinical_data_control$`Record ID`)

LPS_data$age <- clinical_data_full[match(LPS_data$SAMPLE.ID,clinical_data_full$record_id),"s04_age_BL"]%>%unlist()
LPS_data$gender <- clinical_data_full[match(LPS_data$SAMPLE.ID,clinical_data_full$record_id),"s04_gender_BL"]%>%unlist()
LPS_data$bmi <- clinical_data_full[match(LPS_data$SAMPLE.ID,clinical_data_full$record_id),"m02_bmi_BL"]%>%unlist()
LPS_data$hba1c <- clinical_data_full[match(LPS_data$SAMPLE.ID,clinical_data_full$record_id),"b02_hba1c_BL"]%>%unlist()

samples_control <- LPS_data%>%filter(Cohort=="ADIEU")%>%select(SAMPLE.ID)%>%unlist()
LPS_data[which(LPS_data$Cohort=="ADIEU"),"age"] <- clinical_data_control[match(samples_control,clinical_data_control$`Record ID`),"Age"]%>%unlist()
LPS_data[which(LPS_data$Cohort=="ADIEU"),"gender"] <- clinical_data_control[match(samples_control,clinical_data_control$`Record ID`),"Gender"]%>%unlist()%>%recode(
    "Male (0)" = "0",
    "Female (1)" = "1"
  )
LPS_data[which(LPS_data$Cohort=="ADIEU"),"bmi"] <- clinical_data_control[match(samples_control,clinical_data_control$`Record ID`),"Weight"]%>%unlist()/(clinical_data_control[match(samples_control,clinical_data_control$`Record ID`),"Height"]%>%unlist())^2
LPS_data[which(LPS_data$Cohort=="ADIEU"),"hba1c"] <- clinical_data_control[match(samples_control,clinical_data_control$`Record ID`),"Baseline HbA1c result"]%>%unlist()
```

## read cytokine data

```{r}
l <- load("../Data/cytokine_data_ready.RData")
```

# Create long format data
```{r}
LPS_data$newGroup <- LPS_data$group
LPS_data[which(LPS_data$group!="Control"),"newGroup"] <- "TB"

long_data <- LPS_data %>%
  pivot_longer(
    cols = -c(SAMPLE.ID, TIME, Cohort, group,newGroup,age,gender,bmi,hba1c),
    names_to = "Marker",
    values_to = "value"
  ) %>%
  mutate(
    value = as.numeric(as.character(value)),
    TIME= factor(as.character(TIME),levels=c("BL","M6")),
    group = factor(as.character(group),levels=c("Control","Euglycemia","Stress Hyperglycemia","Pre-Diabetes","Newly Diagnosed Diabetes")),
    newGroup = factor(as.character(newGroup),levels=c("Control","TB"))
  ) %>%
  filter(!is.na(value))%>%
  filter(Marker=="LPS")
```

# Table 1 {.tabset}
## Age, years

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  summarise(
    age_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(age,na.rm=TRUE),
      median(age, na.rm = TRUE),
      quantile(age, 0.25, na.rm = TRUE),
      quantile(age, 0.75, na.rm = TRUE)
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(Cohort) %>%
  summarise(
    age_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(age,na.rm=TRUE),
      median(age, na.rm = TRUE),
      quantile(age, 0.25, na.rm = TRUE),
      quantile(age, 0.75, na.rm = TRUE)
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  ggplot(aes(sample = age)) +
  stat_qq() +
  stat_qq_line()

shapiro.test(
  LPS_data %>%
    filter(TIME != "M6") %>%
    pull(age)
)

wilcox.test(age ~ Cohort,
            data = LPS_data %>% filter(TIME != "M6"))
```

## Age group

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    age_group = case_when(
      age < 36 ~ "<36",
      age >= 36 & age < 51 ~ "36-50",
      age >= 51 ~ ">= 51"
    )
  ) %>%
  count(age_group) %>%
  mutate(
    Age_Group = paste0(
      n, " (",
      round(100 * n / sum(n), 1),
      "%)"
    )
  ) %>%
  select(age_group, Age_Group)

LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    age_group = case_when(
      age < 36 ~ "<36",
      age >= 36 & age < 51 ~ "36-50",
      age >= 51 ~ ">= 51"
    )
  ) %>%
  count(Cohort, age_group) %>%
  group_by(Cohort)%>%
  mutate(
    Age_Group = paste0(
      n, " (",
      round(100 * n / sum(n), 1),
      "%)"
    )
  ) %>%
  select(age_group, Age_Group)

tab <- table(
  LPS_data$Cohort,
  cut(
    LPS_data$age,
    breaks = c(-Inf, 36, 51, Inf),
    labels = c("<36", "36-50", ">=51"),
    right = FALSE
  )
)

chisq.test(tab)
```

## Gender

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  summarise(
    Female = paste0(
      sum(gender == "1", na.rm = TRUE),
      " (",
      round(100 * mean(gender == "1", na.rm = TRUE), 1),
      "%)"
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  summarise(
    Male = paste0(
      sum(gender == "0", na.rm = TRUE),
      " (",
      round(100 * mean(gender == "0", na.rm = TRUE), 1),
      "%)"
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(Cohort) %>%
  summarise(
    Male = paste0(
      sum(gender == "0", na.rm = TRUE),
      " (",
      round(100 * mean(gender == "0", na.rm = TRUE), 1),
      "%)"
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(Cohort) %>%
  summarise(
    Female = paste0(
      sum(gender == "1", na.rm = TRUE),
      " (",
      round(100 * mean(gender == "1", na.rm = TRUE), 1),
      "%)"
    )
  )

tab <- table(
  LPS_data$Cohort[LPS_data$TIME != "M6"],
  LPS_data$gender[LPS_data$TIME != "M6"]
)

chisq.test(tab)
```

## BMI, kg/m2

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  summarise(
    bmi_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(bmi,na.rm=TRUE),
      median(bmi, na.rm = TRUE),
      quantile(bmi, 0.25, na.rm = TRUE),
      quantile(bmi, 0.75, na.rm = TRUE)
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(Cohort) %>%
  summarise(
    bmi_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(bmi,na.rm=TRUE),
      median(bmi, na.rm = TRUE),
      quantile(bmi, 0.25, na.rm = TRUE),
      quantile(bmi, 0.75, na.rm = TRUE)
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  ggplot(aes(sample = bmi)) +
  stat_qq() +
  stat_qq_line()

shapiro.test(
  LPS_data %>%
    filter(TIME != "M6") %>%
    pull(bmi)
)

wilcox.test(bmi ~ Cohort,
            data = LPS_data %>% filter(TIME != "M6"))
```

## BMI group

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    bmi_group = case_when(
      bmi < 18.5 ~ "Underweight",
      bmi >= 18.5 & bmi < 25 ~ "Normal",
      bmi >= 25 & bmi < 29.9 ~ "Overweight" , 
      bmi > 30 ~ "Obese"
    )
  ) %>%
  count(bmi_group) %>%
  mutate(
    Bmi_Group = paste0(
      n, " (",
      round(100 * n / sum(n), 1),
      "%)"
    )
  ) %>%
  select(bmi_group, Bmi_Group)

LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    bmi_group = case_when(
      bmi < 18.5 ~ "Underweight",
      bmi >= 18.5 & bmi < 25 ~ "Normal",
      bmi >= 25 & bmi < 29.9 ~ "Overweight" , 
      bmi > 30 ~ "Obese"
    )
  ) %>%
  count(Cohort,bmi_group) %>%
  group_by(Cohort)%>%
  mutate(
    Bmi_Group = paste0(
      n, " (",
      round(100 * n / sum(n), 1),
      "%)"
    )
  ) %>%
  select(Cohort,bmi_group, Bmi_Group)

tab <- table(
  LPS_data$Cohort,
  cut(
    LPS_data$bmi,
    breaks = c(-Inf, 18.5, 25, 29.9, Inf),
    labels = c("Underweight", "Normal", "Overweight","Obese"),
    right = FALSE
  )
)

fisher.test(tab)
```

## HbA1c

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  summarise(
    hba1c_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(hba1c,na.rm=TRUE),
      median(hba1c, na.rm = TRUE),
      quantile(hba1c, 0.25, na.rm = TRUE),
      quantile(hba1c, 0.75, na.rm = TRUE)
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(Cohort) %>%
  summarise(
    hba1c_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(hba1c,na.rm=TRUE),
      median(hba1c, na.rm = TRUE),
      quantile(hba1c, 0.25, na.rm = TRUE),
      quantile(hba1c, 0.75, na.rm = TRUE)
    )
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  ggplot(aes(sample = hba1c)) +
  stat_qq() +
  stat_qq_line()

shapiro.test(
  LPS_data %>%
    filter(TIME != "M6") %>%
    pull(hba1c)
)

wilcox.test(hba1c ~ Cohort,
            data = LPS_data %>% filter(TIME != "M6"))
```

## HbA1c group

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    hba1c_group = case_when(
      hba1c < 5.7 ~ "Normal",
      hba1c >= 5.7 & hba1c < 6.5 ~ "Pre-diabetes",
      hba1c >= 6.5  ~ "Diabetes" 
    )
  ) %>%
  count(hba1c_group) %>%
  mutate(
    Hba1c_Group = paste0(
      n, " (",
      round(100 * n / sum(n), 1),
      "%)"
    )
  ) %>%
  select(hba1c_group, Hba1c_Group)

LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    hba1c_group = case_when(
      hba1c < 5.7 ~ "Normal",
      hba1c >= 5.7 & hba1c < 6.5 ~ "Pre-diabetes",
      hba1c >= 6.5  ~ "Diabetes" 
    )
  ) %>%
  count(Cohort,hba1c_group) %>%
  group_by(Cohort)%>%
  mutate(
    Hba1c_Group = paste0(
      n, " (",
      round(100 * n / sum(n), 1),
      "%)"
    )
  ) %>%
  select(Cohort,hba1c_group, Hba1c_Group)

tab <- table(
  LPS_data$Cohort,
  cut(
    LPS_data$hba1c,
    breaks = c(-Inf, 5.7, 6.5, Inf),
    labels = c("Normal", "Pre-diabetes","Diabetes"),
    right = FALSE
  )
)

fisher.test(tab)
```

# Table 2 {.tabset}
## Association LPS with group

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(newGroup) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

wilcox.test(LPS ~ newGroup,
            data = LPS_data %>%
  filter(TIME != "M6"))
```

## Association LPS with gender

```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(gender) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )


LPS_data %>%
  filter(TIME != "M6") %>%
  group_by(Cohort, gender) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

wilcox.test(LPS ~ gender,
            data = LPS_data %>%
  filter(TIME != "M6"))

wilcox.test(LPS ~ gender,
            data = LPS_data %>%
  filter(TIME != "M6", Cohort=="MIDST"))

wilcox.test(LPS ~ gender,
            data = LPS_data %>%
  filter(TIME != "M6", Cohort=="ADIEU"))
```

## Association LPS with age group
```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    age_group = case_when(
      age < 36 ~ "<36",
      age >= 36 & age < 51 ~ "36-50",
      age >= 51 ~ ">=51"
    )
  ) %>%
  group_by(age_group) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

dat <- LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    age_group = case_when(
      age < 36 ~ "<36",
      age >= 36 & age < 51 ~ "36-50",
      age >= 51 ~ ">=51"
    )
  )
kruskal.test(LPS ~ age_group, data = dat)

LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    age_group = case_when(
      age < 36 ~ "<36",
      age >= 36 & age < 51 ~ "36-50",
      age >= 51 ~ ">=51"
    )
  ) %>%
  group_by(Cohort, age_group) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    age_group = case_when(
      age < 36 ~ "<36",
      age >= 36 & age < 51 ~ "36-50",
      age >= 51 ~ ">=51"
    )
  ) %>%
  group_by(Cohort) %>%
  summarise(
    p_value = kruskal.test(LPS ~ age_group)$p.value
  )
```

## Association LPS with BMI group
```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    bmi_group = case_when(
      bmi < 18.5 ~ "Underweight",
      bmi >= 18.5 & bmi < 25 ~ "Normal",
      bmi >= 25 & bmi < 29.9 ~ "Overweight" , 
      bmi > 30 ~ "Obese"
    )
  ) %>%
  group_by(bmi_group) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

dat <- LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    bmi_group = case_when(
      bmi < 18.5 ~ "Underweight",
      bmi >= 18.5 & bmi < 25 ~ "Normal",
      bmi >= 25 & bmi < 29.9 ~ "Overweight" , 
      bmi > 30 ~ "Obese"
    )
  )
kruskal.test(LPS ~ bmi_group, data = dat)

LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    bmi_group = case_when(
      bmi < 18.5 ~ "Underweight",
      bmi >= 18.5 & bmi < 25 ~ "Normal",
      bmi >= 25 & bmi < 29.9 ~ "Overweight" , 
      bmi > 30 ~ "Obese"
    )
  ) %>%
  group_by(Cohort, bmi_group) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

LPS_data %>%
  filter(TIME != "M6", Cohort=="MIDST") %>%
  mutate(
    bmi_group = case_when(
      bmi < 18.5 ~ "Underweight",
      bmi >= 18.5 & bmi < 25 ~ "Normal",
      bmi >= 25 & bmi < 29.9 ~ "Overweight" , 
      bmi > 30 ~ "Obese"
    )
  ) %>%
  summarise(
    p_value = kruskal.test(LPS ~ bmi_group)$p.value
  )

LPS_data %>%
  filter(TIME != "M6", Cohort=="ADIEU") %>%
  mutate(
    bmi_group = case_when(
      bmi >= 18.5 & bmi < 25 ~ "Normal",
      bmi >= 25 & bmi < 29.9 ~ "Overweight" , 
      bmi > 30 ~ "Obese"
    )
  ) %>%
  summarise(
    p_value = kruskal.test(LPS ~ bmi_group)$p.value
  )
```

## Association LPS with HbA1c group
```{r}
LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
     hba1c_group = case_when(
      hba1c < 5.7 ~ "Normal",
      hba1c >= 5.7 & hba1c < 6.5 ~ "Pre-diabetes",
      hba1c >= 6.5  ~ "Diabetes" 
    )
  ) %>%
  group_by(hba1c_group) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

dat <- LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
     hba1c_group = case_when(
      hba1c < 5.7 ~ "Normal",
      hba1c >= 5.7 & hba1c < 6.5 ~ "Pre-diabetes",
      hba1c >= 6.5  ~ "Diabetes" 
    )
  )
kruskal.test(LPS ~ hba1c_group, data = dat)

LPS_data %>%
  filter(TIME != "M6") %>%
  mutate(
    hba1c_group = case_when(
      hba1c < 5.7 ~ "Normal",
      hba1c >= 5.7 & hba1c < 6.5 ~ "Pre-diabetes",
      hba1c >= 6.5  ~ "Diabetes" 
    )
  ) %>%
  group_by(Cohort, hba1c_group) %>%
  summarise(
    LPS_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(LPS, na.rm = TRUE),
      median(LPS, na.rm = TRUE),
      quantile(LPS, 0.25, na.rm = TRUE),
      quantile(LPS, 0.75, na.rm = TRUE)
    ),
    .groups = "drop"
  )

LPS_data %>%
  filter(TIME != "M6", Cohort=="MIDST") %>%
  mutate(
     hba1c_group = case_when(
      hba1c < 5.7 ~ "Normal",
      hba1c >= 5.7 & hba1c < 6.5 ~ "Pre-diabetes",
      hba1c >= 6.5  ~ "Diabetes" 
    )
  ) %>%
  summarise(
    p_value = kruskal.test(LPS ~ hba1c_group)$p.value
  )
```

# Figure 1 {.tabset}
## LPS levels in TB at BL and M6 and controls

```{r}
input_data <- long_data

input_data %>%
  group_by(Cohort, TIME) %>%
  summarise(
    age_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(value,na.rm=TRUE),
      median(value, na.rm = TRUE),
      quantile(value, 0.25, na.rm = TRUE),
      quantile(value, 0.75, na.rm = TRUE)
    )
  )

input_data <- long_data
input_data %>%
  group_by(Cohort, group, TIME) %>%
  summarise(
    age_summary = sprintf(
      "%.2f %.2f (%.2f–%.2f)",
      mean(value,na.rm=TRUE),
      median(value, na.rm = TRUE),
      quantile(value, 0.25, na.rm = TRUE),
      quantile(value, 0.75, na.rm = TRUE)
    )
  )

input_data %>%
  filter(TIME != "M6") %>%
  ggplot(aes(sample = value)) +
  stat_qq() +
  stat_qq_line()

shapiro.test(
  input_data %>%
    filter(TIME != "M6") %>%
    pull(value)
)

wilcox.test(value ~ Cohort,
            data = input_data %>% filter(TIME != "M6"))
```

## Plot for LPS in TB vs. control

```{r}
input_data <- long_data
input_data$newTIME <- paste0(input_data$TIME,input_data$newGroup)
input_data$newTIME <- factor(input_data$newTIME,levels=c("BLTB","M6TB","BLControl"))
input_data$value <- log10(input_data$value)

input_data %>%
  group_by(newGroup,TIME) %>%
  summarise(
    W = shapiro.test(value)$statistic,
    p.value = shapiro.test(value)$p.value
  )

Figure1 <- ggplot(input_data, aes(x = newTIME, y = value, fill=newGroup)) +
  
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  
  geom_jitter(
    aes(color = newGroup),
    position = position_jitter(width = 0.1),
    size = 3,
    alpha = 0.8
  ) +
  
 scale_fill_manual(values = c(
    "Control" = "black",
    "TB" = "darkred"
  ),labels=c(
    "Control"="Asymptomatic non-TB controls",
    "TB" = "TB")) +
  
  scale_color_manual(values = c(
    "Control" = "black",
     "TB" = "darkred"
  ), labels=c(
     "Control"="Asymptomatic non-TB controls",
     "TB" = "TB")) +
  
  stat_compare_means(
    comparisons = list(c("BLTB", "BLControl"),c("BLTB", "M6TB"),c("M6TB", "BLControl")),
    method = "wilcox.test",
    label = "p.signif",
    hide.ns = FALSE,
   step.increase = 0.06,
   size = 6
  ) +
  scale_x_discrete(labels = c(
    "BLTB" = "PTB at Baseline",
    "M6TB" = "PTB at Month 6",
    "BLControl" = "non-TB controls"
    
  )) +
  theme_classic() +
  theme(legend.position = "none")+
  labs(x = "", y =expression(log[10]("LPS (EU/ml)")), color = "",fill="")+
  theme(
  axis.text = element_text(size=26),
  axis.title = element_text(size=26),
  legend.position = "none",
  axis.line.x.top = element_blank(),
  axis.text.x.top = element_blank(),
  axis.ticks.x.top = element_blank(),
  axis.title.x.top = element_blank()
)
```

## Make Figure 1

```{r}
pdf("../Results/New Manuscript/Figure 1.pdf",width=11,height = 11)
Figure1
dev.off()
```



# Table 3: Regression analysis {.tabset}
### Association between LPS and TB at baseline

```{r}
input_data <- long_data%>%
  filter(TIME=="BL")
model <- lm(log10(value) ~ newGroup + age + gender + bmi ,data = input_data)
summary_model <- summary(model)

tbl <- data.frame(
    outcome = "LPS",
    term = rownames(summary_model$coefficients),
    estimate = summary_model$coefficients[,1],
    std.error = summary_model$coefficients[,2],
    p.value = summary_model$coefficients[,4]
  )%>%filter(term!="(Intercept)")%>%
  gt()%>%
  tab_header(
    title = "Linear Regression",
    subtitle = "Model at BL: LPS ~ newGroup + age + gender + bmi"
  )
```

### Association between LPS and TB at month 6

```{r}
input_data <- long_data%>%
  filter(TIME=="M6"|newGroup=="Control")
model <- lm(log10(value) ~ newGroup + age + gender + bmi ,data = input_data)
summary_model <- summary(model)

tbl <- data.frame(
    outcome = "LPS",
    term = rownames(summary_model$coefficients),
    estimate = summary_model$coefficients[,1],
    std.error = summary_model$coefficients[,2],
    p.value = summary_model$coefficients[,4]
  )%>%filter(term!="(Intercept)")%>%
  gt()%>%
  tab_header(
    title = "Linear Regression",
    subtitle = "Model at M6: LPS ~ newGroup + age + gender + bmi"
  )
```

# Figure 2 {.tabset}
## Panel a: not broken y axis only for BL 
```{r}
input_data <- long_data%>%
  filter(TIME == "BL")%>%
  filter(group=="Euglycemia"|group=="Stress Hyperglycemia"|group=="Control")
input_data$value <- log10(input_data$value)

input_data %>%
  group_by(group) %>%
  summarise(
    W = shapiro.test(value)$statistic,
    p.value = shapiro.test(value)$p.value
  )

Figure2_a <- input_data %>%
  ggplot(aes(x = group, y = value, fill = group)) +
  
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(aes(color = group),
              position = position_jitter(width = 0.1),
              size = 3, alpha = 0.8) +
  
  stat_compare_means(
    comparisons = list(
      c("Stress Hyperglycemia", "Euglycemia"),
      c("Stress Hyperglycemia", "Control"),
      c("Euglycemia", "Control")
    ),
    method = "wilcox.test",
    label = "p.signif",
    size=6
  ) +
  scale_fill_manual(values = c(
    "Control" = "black",
    "Diabetes" = "#984EA3",
    "Stress Hyperglycemia" = "#3abbcc",
    "Euglycemia" = "#ca6332"
  ),labels=c(
    "Control"="Asymptomatic non-TB controls",
    "Diabetes" = "TB-diabetes",
    "Stress Hyperglycemia" = "TB-stress hyperglycemia",
    "Euglycemia" = "TB-euglycemia")) +
  
  scale_color_manual(values = c(
    "Control" = "black",
    "Diabetes" = "#984EA3",
    "Stress Hyperglycemia" = "#3abbcc",
    "Euglycemia" = "#ca6332"
  ), labels=c(
     "Control"="non-TB controls",
    "Diabetes" = "TB-diabetes",
    "Stress Hyperglycemia" = "TB-stress hyperglycemia",
    "Euglycemia" = "TB-euglycemia")) +
  
   scale_x_discrete(labels = c(
    "Control" = "non-TB controls",
    "Diabetes" = "TB-diabetes",
    "Stress Hyperglycemia" = "TB-stress hyperglycemia",
    "Euglycemia" = "TB-euglycemia"
  )) +
  labs(title = "Baseline",y=expression(log[10]("LPS (EU/ml)")),fill="",color="",size=3) +
  theme_classic()+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text = element_text(size=23),
        plot.title = element_text(size = 22, face = "bold"),
  axis.title = element_text(size=26))

pdf("../Results/New Manuscript/Figure 2-a.pdf",height = 11, width = 11)
Figure2_a
dev.off()
```

## Panel b: not broken y axis only for M6

```{r}
input_data <- long_data%>%
  filter(TIME == "M6")%>%
  filter(group=="Euglycemia"|group=="Stress Hyperglycemia")%>%
  bind_rows(long_data%>%
  filter(TIME == "BL",group=="Control")%>%mutate(TIME="M6"))

input_data$value <- log10(input_data$value)

input_data %>%
  group_by(group) %>%
  summarise(
    W = shapiro.test(value)$statistic,
    p.value = shapiro.test(value)$p.value
  )

Figure2_b <- input_data %>%
  ggplot(aes(x = group, y = value, fill = group)) +
  
  geom_boxplot(alpha = 0.6, outlier.shape = NA) +
  geom_jitter(aes(color = group),
              position = position_jitter(width = 0.1),
              size = 3, alpha = 0.8) +
  
  stat_compare_means(
    comparisons = list(
      c("Stress Hyperglycemia", "Euglycemia"),
      c("Stress Hyperglycemia", "Control"),
      c("Euglycemia", "Control")
    ),
    method = "wilcox.test",
    label = "p.signif",
    size=6
  ) +
  scale_fill_manual(values = c(
    "Control" = "black",
    "Diabetes" = "#984EA3",
    "Stress Hyperglycemia" = "#3abbcc",
    "Euglycemia" = "#ca6332"
  ),labels=c(
    "Control"="non-TB controls",
    "Diabetes" = "TB-diabetes",
    "Stress Hyperglycemia" = "TB-stress hyperglycemia",
    "Euglycemia" = "TB-euglycemia")) +
  
  scale_color_manual(values = c(
    "Control" = "black",
    "Diabetes" = "#984EA3",
    "Stress Hyperglycemia" = "#3abbcc",
    "Euglycemia" = "#ca6332"
  ), labels=c(
     "Control"="non-TB controls",
    "Diabetes" = "TB-diabetes",
    "Stress Hyperglycemia" = "TB-stress hyperglycemia",
    "Euglycemia" = "TB-euglycemia")) +
   scale_x_discrete(labels = c(
    "Control" = "non-TB controls",
    "Diabetes" = "TB-diabetes",
    "Stress Hyperglycemia" = "TB-stress hyperglycemia",
    "Euglycemia" = "TB-euglycemia"
  )) +
  labs(title = "Month 6",y=expression(log[10]("LPS (EU/ml)")),fill="",color="") +
  theme_classic()+
  theme(legend.position = "none",
        axis.title.x = element_blank(),
       # axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text = element_text(size=23),
        plot.title = element_text(size = 22, face = "bold"),
  axis.title = element_text(size=26))


pdf("../Results/New Manuscript/Figure 2-b.pdf",height = 11, width = 11)
Figure2_b
dev.off()
```

## Panel c : Changes in LPS from BL to M6 in each glycemic group

```{r}
input_data <- long_data%>%
  filter(group=="Euglycemia"|group=="Stress Hyperglycemia")

input_data$value <- log10(input_data$value)

stat.test <- input_data %>%
  group_by(group) %>%
  summarise(
    p = wilcox.test(value[TIME == "BL"],
                    value[TIME == "M6"],paired=TRUE)$p.value,
    y.position = max(value, na.rm = TRUE) * 1.2,
    .groups = "drop"
  ) %>%
  mutate(
    p.signif = case_when(
      p <= 0.001 ~ "***",
      p <= 0.01  ~ "**",
      p <= 0.05  ~ "*",
      TRUE       ~ "ns"
    ),
    group1 = "BL",
    group2 = "M6"
  )

Figure2_c <- ggplot(input_data, aes(x = TIME, y = value, group = SAMPLE.ID)) +
  
  geom_line(alpha = 0.4, color = "gray70") +
  geom_point(aes(color = group), size = 3, alpha = 0.9) +
  
  facet_wrap(~group, scales = "free_y", nrow = 1,
             labeller = as_labeller(c(
               "Stress Hyperglycemia" = "TB-stress hyperglycemia",
               "Euglycemia" = "TB-euglycemia"
             ))) +
  
  scale_color_manual(values = c(
    "Stress Hyperglycemia" = "#3abbcc",
    "Euglycemia" = "#ca6332"
  )) +
  
  stat_pvalue_manual(
    stat.test,
    label = "p.signif",
    xmin = "group1",
    xmax = "group2",
    y.position = "y.position",
    inherit.aes = FALSE,
    size=6
  ) +
  
  scale_x_discrete(labels = c(
    "BL" = "Baseline",
    "M6" = "Month 6"
  )) +
  theme_classic() +
  theme( legend.position = "none",
        axis.title.x = element_blank(),
       # axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text = element_text(size=23),
        plot.title = element_text(size = 22, face = "bold"),
  axis.title = element_text(size=26),
  strip.text = element_text(size = 22, face = "bold"))+
  labs(x = "", y = expression(log[10]("LPS (EU/ml)")), color = "")

pdf("../Results/New Manuscript/Figure 2-c.pdf",height = 11, width = 9)
Figure2_c
dev.off()
```

## Make Figure 2

```{r, eval=FALSE}
Figure2_a <- Figure2_a + coord_cartesian(ylim = c(-2, 3))
Figure2_b <- Figure2_b + coord_cartesian(ylim = c(-2, 3))
Figure2_c <- Figure2_c + coord_cartesian(ylim = c(-2, 3))

pdf("../Results/New Manuscript/Figure 2.pdf",height = 11, width = 32)
plot_grid(Figure2_a,Figure2_b,Figure2_c,labels=c("a","b","c"),label_size = 35,rel_widths = c(1,1,1),align = "h",axis = "tb",nrow = 1)
dev.off()
```

# Table 3: Regression analysis {.tabset}
## Association between glycemic status and LPS at baseline

```{r}
input_data <- long_data%>%
  filter(TIME=="BL")%>%
  filter(group=="Euglycemia"|group=="Stress Hyperglycemia")

model <- lm(log10(value) ~ group + age + gender + bmi ,data = input_data)
summary_model <- summary(model)

tbl <- data.frame(
    outcome = "LPS",
    term = rownames(summary_model$coefficients),
    estimate = summary_model$coefficients[,1],
    std.error = summary_model$coefficients[,2],
    p.value = summary_model$coefficients[,4]
  )%>%filter(term!="(Intercept)")%>%
  gt()%>%
  tab_header(
    title = "Linear Regression",
    subtitle = "Model at BL: LPS ~ group + age + gender + bmi"
  )
```

## Association between glycemic status and LPS at month 6

```{r}
input_data <- long_data%>%
  filter(TIME=="M6")%>%
  filter(group=="Euglycemia"|group=="Stress Hyperglycemia")

model <- lm(log10(value) ~ group + age + gender + bmi ,data = input_data)
summary_model <- summary(model)

tbl <- data.frame(
    outcome = "LPS",
    term = rownames(summary_model$coefficients),
    estimate = summary_model$coefficients[,1],
    std.error = summary_model$coefficients[,2],
    p.value = summary_model$coefficients[,4]
  )%>%filter(term!="(Intercept)")%>%
  gt()%>%
  tab_header(
    title = "Linear Regression",
    subtitle = "Model at M6: LPS ~ group + age + gender + bmi"
  )
```

# Figure 3 {.tabset}
## Make data for LPS at baseline and clinical outcomes

```{r}
clinical_data_full <- read_excel("../Data/MIDST_ZMR21.xlsx")

LPS_data_BL_MIDST <- LPS_data%>%
  filter(TIME=="BL",Cohort=="MIDST")

LPS_data_M6_MIDST <- LPS_data%>%
  filter(TIME=="M6",Cohort=="MIDST")

LPS_data_BL_MIDST$deltaLPS <- log10(LPS_data_M6_MIDST[match(LPS_data_BL_MIDST$SAMPLE.ID,LPS_data_M6_MIDST$SAMPLE.ID),"LPS"])-log10(LPS_data_BL_MIDST[,"LPS"])

clinical_data_full <- clinical_data_full[match(LPS_data_BL_MIDST$SAMPLE.ID,clinical_data_full$record_id),]

clinical_data_full$weightBL <- clinical_data_full$m02_weight_BL
clinical_data_full$bmiBL <- clinical_data_full$m02_bmi_BL
clinical_data_full$homairBL <- clinical_data_full$u01_homair_BL

clinical_data_full$delta_weightM2 <- clinical_data_full$m02_weight_M2 - clinical_data_full$m02_weight_BL
clinical_data_full$delta_bmiM2 <- clinical_data_full$m02_bmi_M2 - clinical_data_full$m02_bmi_BL
clinical_data_full$delta_homairM2 <- clinical_data_full$u01_homair_M2 - clinical_data_full$u01_homair_BL

clinical_data_full$prop_delta_weightM2 <- (clinical_data_full$m02_weight_M2 - clinical_data_full$m02_weight_BL)/clinical_data_full$m02_weight_BL
clinical_data_full$prop_delta_bmiM2 <- (clinical_data_full$m02_bmi_M2 - clinical_data_full$m02_bmi_BL)/clinical_data_full$m02_bmi_BL
clinical_data_full$prop_delta_homairM2 <- (clinical_data_full$u01_homair_M2 - clinical_data_full$u01_homair_BL)/clinical_data_full$u01_homair_BL

clinical_data_full$delta_weightM6 <- clinical_data_full$m02_weight_M6 - clinical_data_full$m02_weight_BL
clinical_data_full$delta_bmiM6 <- clinical_data_full$m02_bmi_M6 - clinical_data_full$m02_bmi_BL
clinical_data_full$delta_homairM6 <- clinical_data_full$u01_homair_M6 - clinical_data_full$u01_homair_BL

clinical_data_full$prop_delta_weightM6 <- (clinical_data_full$m02_weight_M6 - clinical_data_full$m02_weight_BL)/clinical_data_full$m02_weight_BL
clinical_data_full$prop_delta_bmiM6 <- (clinical_data_full$m02_bmi_M6 - clinical_data_full$m02_bmi_BL)/clinical_data_full$m02_bmi_BL
clinical_data_full$prop_delta_homairM6 <- (clinical_data_full$u01_homair_M6 - clinical_data_full$u01_homair_BL)/clinical_data_full$u01_homair_BL
clinical_data_full$prop_delta_homairM6 <- (clinical_data_full$u01_homair_M6 - clinical_data_full$u01_homair_BL)

input_data <- data.frame(
  Sample_ID = LPS_data_BL_MIDST$SAMPLE.ID,
  LPS = LPS_data_BL_MIDST$LPS,
  deltaLPS = LPS_data_BL_MIDST$deltaLPS,
  age = clinical_data_full$s04_age_BL,
  gender = clinical_data_full$s04_gender_BL,
  weightBL = clinical_data_full$weightBL,
  bmiBL = clinical_data_full$bmiBL,
  homairBL = clinical_data_full$homair_BL,
  delta_weightM2 = clinical_data_full$delta_weightM2 ,
  delta_bmiM2 = clinical_data_full$delta_bmiM2,
  delta_homairM2 = clinical_data_full$delta_homairM2,
  prop_delta_weightM2 = clinical_data_full$prop_delta_weightM2 ,
  prop_delta_bmiM2 = clinical_data_full$prop_delta_bmiM2,
  prop_delta_homairM2 = clinical_data_full$prop_delta_homairM2,
  delta_weightM6 = clinical_data_full$delta_weightM6 ,
  delta_bmiM6 = clinical_data_full$delta_bmiM6,
  delta_homairM6 = clinical_data_full$delta_homairM6,
  prop_delta_weightM6 = clinical_data_full$prop_delta_weightM6 ,
  prop_delta_bmiM6 = clinical_data_full$prop_delta_bmiM6,
  prop_delta_homairM6 = clinical_data_full$prop_delta_homairM6,
  rel_weight_change_BL_M2=clinical_data_full$rel_weight_change_BL_M2,
  rel_weight_change_BL_M6=clinical_data_full$rel_weight_change_BL_M6,
  Group = LPS_data_BL_MIDST$group
)

input_data_BL <- input_data%>%
  select(Sample_ID,LPS,weightBL,bmiBL,homairBL,age, gender, Group)%>%
  mutate(time="BL")%>%
  rename(Weight=weightBL,BMI=bmiBL,HOMA_IR=homairBL)
input_data_BL$LPS <- log10(input_data_BL$LPS)

input_data_BL_change <- input_data%>%
  select(Sample_ID,LPS,rel_weight_change_BL_M2,rel_weight_change_BL_M6,age, gender,Group)%>%
  mutate(time="BL")%>%
  rename(WeightBLM2=rel_weight_change_BL_M2,WeightBLM6=rel_weight_change_BL_M6)
input_data_BL_change$LPS <- log10(input_data_BL_change$LPS)

input_data_BL_change_2 <- input_data%>%
  select(Sample_ID,deltaLPS,rel_weight_change_BL_M6,prop_delta_bmiM6, prop_delta_homairM6, age, gender, Group)%>%
  mutate(time="BL")%>%
  rename(Weight=rel_weight_change_BL_M6,BMI=prop_delta_bmiM6,HOMA_IR=prop_delta_homairM6)
```

## Make data for LPS at M6 and clinical outcomes

```{r}
clinical_data_full <- read_excel("../Data/MIDST_ZMR21.xlsx")

LPS_data_M6_MIDST <- LPS_data%>%
  filter(TIME=="M6",Cohort=="MIDST")

clinical_data_full <- clinical_data_full[match(LPS_data_M6_MIDST$SAMPLE.ID,clinical_data_full$record_id),]

clinical_data_full$weightM6 <- clinical_data_full$m02_weight_M6
clinical_data_full$bmiM6 <- clinical_data_full$m02_bmi_M6
clinical_data_full$homairM6 <- clinical_data_full$u01_homair_M6

clinical_data_full$prop_delta_bmiM6 <- (clinical_data_full$m02_bmi_M6 - clinical_data_full$m02_bmi_BL)/clinical_data_full$m02_bmi_BL
clinical_data_full$prop_delta_homairM6 <- (clinical_data_full$u01_homair_M6 - clinical_data_full$u01_homair_BL)/clinical_data_full$u01_homair_BL
clinical_data_full$prop_delta_homairM6 <- (clinical_data_full$u01_homair_M6 - clinical_data_full$u01_homair_BL)


input_data <- data.frame(
  Sample_ID = LPS_data_M6_MIDST$SAMPLE.ID,
  LPS = LPS_data_M6_MIDST$LPS,
  age = clinical_data_full$s04_age_BL,
  gender = clinical_data_full$s04_gender_BL,
  weightM6 = clinical_data_full$weightM6,
  bmiM6 = clinical_data_full$bmiM6,
  homairM6 = clinical_data_full$homairM6,
  prop_delta_bmiM6 = clinical_data_full$prop_delta_bmiM6,
  prop_delta_homairM6 = clinical_data_full$prop_delta_homairM6,
  rel_weight_change_BL_M6=clinical_data_full$rel_weight_change_BL_M6,
  Group = LPS_data_M6_MIDST$group
)

input_data_M6 <- input_data%>%
  select(Sample_ID,LPS,weightM6,bmiM6,homairM6,age, gender,prop_delta_bmiM6,prop_delta_homairM6,rel_weight_change_BL_M6, Group)%>%
  mutate(time="M6")%>%
  rename(Weight=weightM6,BMI=bmiM6,HOMA_IR=homairM6,deltaWeight=rel_weight_change_BL_M6,deltaBMI=prop_delta_bmiM6,deltaHOMA_IR=prop_delta_homairM6)
input_data_M6$LPS <- log10(input_data_M6$LPS)
```

## Make Figure 3
```{r}
input_data <- input_data_BL%>%
  bind_rows(input_data_M6)%>%
  pivot_longer(
    cols = c(Weight,BMI,HOMA_IR),
    names_to = "marker",
    values_to = "value"
  )%>%
  mutate(
    time = recode(time,
                  "BL" = "Baseline",
                  "M6" = "Month 6")
  )

input_data$marker <- factor(input_data$marker,levels=c("Weight","BMI","HOMA_IR"))
Figure3_a <- plot_LPS_marker_withoutfacet(input_data, m="Weight",t="Baseline",method="spearman")
Figure3_b <- plot_LPS_marker_withoutfacet(input_data, m="BMI",t="Baseline",method="spearman")
Figure3_c <- plot_LPS_marker_withoutfacet(input_data, m="HOMA_IR",t="Baseline",method="spearman")
Figure3_d <- plot_LPS_marker_withoutfacet(input_data, m="Weight",t="Month 6",method="spearman")
Figure3_e <- plot_LPS_marker_withoutfacet(input_data, m="BMI",t="Month 6",method="spearman")
Figure3_f <- plot_LPS_marker_withoutfacet(input_data, m="HOMA_IR",t="Month 6",method="spearman")

pdf("../Results/New Manuscript/Figure 3.pdf",height =12, width = 18)
plot_grid(Figure3_b,Figure3_c,Figure3_e,Figure3_f,labels=c("a","b","c","d"),label_size = 20,rel_widths = c(1,1),align = "h",axis = "tb",nrow = 2)
dev.off()
```

## Make Supplementary Figure 1

```{r}
input_data_BL_change_2_long <- input_data_BL_change_2%>%
  pivot_longer(
    cols = c(Weight,BMI,HOMA_IR),
    names_to = "marker",
    values_to = "value"
  )%>%
  mutate(
    time = recode(time,
                  "BL" = "Baseline",
                  "M6" = "Month 6")
  )
input_data_BL_change_2_long$marker <- factor(input_data_BL_change_2_long$marker,levels=c("Weight","BMI","HOMA_IR"))
Figure3_g <- plot_deltaLPS_deltamarker_withoutfacet(input_data_BL_change_2_long,"Weight",method="spearman")
Figure3_h <- plot_deltaLPS_deltamarker_withoutfacet(input_data_BL_change_2_long,"BMI",method="spearman")
Figure3_i <- plot_deltaLPS_deltamarker_withoutfacet(input_data_BL_change_2_long,"HOMA_IR",method="spearman")
pdf("../Results/New Manuscript/Supplementary Figure 1.pdf",height = 10, width = 18)
plot_grid(Figure3_h,Figure3_i,labels=c("a","b"),label_size = 20,rel_widths = c(1,1,1),align = "h",axis = "tb",nrow = 1)
dev.off()
```

# Table 4: Regression analysis {.tabset}
## Association between LPS and clinical outcomes at baseline 
```{r}
outcomes <- c(
  "Weight",
  "BMI",
  "HOMA_IR"
)

results <- setNames(
  lapply(outcomes, function(var) {
    lm(as.formula(paste("LPS ~ ", var,"+ age + gender + Group")), data = input_data_BL)
  }),
  outcomes
)

all_results <- lapply(names(results), function(v) {
  summary_model <- summary(results[[v]])
  
  data.frame(
    outcome = "LPS",
    term = rownames(summary_model$coefficients),
    estimate = summary_model$coefficients[,1],
    std.error = summary_model$coefficients[,2],
    p.value = summary_model$coefficients[,4]
  )
})

all_results <- do.call(rbind, all_results)
tbl <- all_results%>%filter(term!="(Intercept)")%>%
  gt()%>%
  tab_header(
    title = "Linear Regression",
    subtitle = "Model at BL: LPS ~ var + age + gender + Group"
  )

table3 <- data.frame(
  model = character(),
  term = character(),
  estimate = numeric(),
  p_value = numeric(),
  stringsAsFactors = FALSE
)
table3 <- table3 %>%
  bind_rows(
    data.frame(
      model = "LPS ~ Weight + age + gender + Group",
      term = "Weight",
      estimate = all_results["Weight","estimate"],
      p_value = all_results["Weight","p.value"]
    )
  )%>%
  bind_rows(
    data.frame(
      model = "LPS ~ BMI + age + gender + Group",
      term = "BMI",
      estimate = all_results["BMI", "estimate"],
      p_value = all_results["BMI", "p.value"]
    )
  )%>%
  bind_rows(
    data.frame(
      model = "LPS ~ HOMA_IR + age + gender + Group",
      term = "HOMA_IR",
      estimate = all_results["HOMA_IR", "estimate"],
      p_value = all_results["HOMA_IR", "p.value"]
    )
  )
```

## Association between LPS and clinical outcomes at month 6

```{r}
outcomes <- c(
  "Weight",
  "BMI",
  "HOMA_IR"
)

results <- setNames(
  lapply(outcomes, function(var) {
    lm(as.formula(paste("LPS ~ ",var, "+ age + gender + Group")), data = input_data_M6)
  }),
  outcomes
)

all_results <- lapply(names(results), function(v) {
  summary_model <- summary(results[[v]])
  
  data.frame(
    outcome = "LPS",
    term = rownames(summary_model$coefficients),
    estimate = summary_model$coefficients[,1],
    std.error = summary_model$coefficients[,2],
    p.value = summary_model$coefficients[,4]
  )
})

all_results <- do.call(rbind, all_results)
tbl <- all_results%>%filter(term!="(Intercept)")%>%
  gt()%>%
  tab_header(
    title = "Linear Regression",
    subtitle = "Model at M6: LPS ~ var + age + gender + Group"
  )

table3 <- table3 %>%
  bind_rows(
    data.frame(
      model = "LPS ~ Weight + age + gender + Group",
      term = "Weight",
      estimate = all_results["Weight","estimate"],
      p_value = all_results["Weight","p.value"]
    )
  )%>%
  bind_rows(
    data.frame(
      model = "LPS ~ BMI + age + gender + Group",
      term = "BMI",
      estimate = all_results["BMI", "estimate"],
      p_value = all_results["BMI", "p.value"]
    )
  )%>%
  bind_rows(
    data.frame(
      model = "LPS ~ HOMA_IR + age + gender + Group",
      term = "HOMA_IR",
      estimate = all_results["HOMA_IR", "estimate"],
      p_value = all_results["HOMA_IR", "p.value"]
    )
  )

```

# Figure 4 {.tabset}
## All TB {.tabset}
### Correlation between LPS at baseline and cytokines at baseline

```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL", Cohort=="MIDST")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL",newGroup!="Control")
LPS_data_BL <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_BL <- cytokine_data_BL[match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
n_BLBL_TB <- nrow(cytokine_data_BL)
data_1 <- LPS_data_BL%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_BL%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_BLBL_TB <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p = df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_BLBL <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_BLBL <- df_sig$Cytokine%>%unique()
df_sig
```


### Correlation between LPS at month 6 and cytokines at month 6
```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6")
LPS_data_M6 <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_M6 <- cytokine_data_M6[match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
n_M6M6_TB <- nrow(cytokine_data_M6)
data_1 <- LPS_data_M6%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_M6%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_M6M6_TB <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p = df_long$p.value)


df_sig <- df_long%>%filter(p.value<0.05)

links_M6M6 <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_M6M6 <- df_sig$Cytokine%>%unique()
df_sig
```

### Correlation between LPS at baseline and cytokine at month 6
```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6")
LPS_data_BL_M6 <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_BL_M6 <- cytokine_data_M6[match(LPS_data_BL_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
n_BLM6_TB <- nrow(cytokine_data_BL_M6)
data_1 <- LPS_data_BL_M6%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_BL_M6%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_BLM6_TB <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p = df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_BLM6 <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_BLM6 <- df_sig$Cytokine%>%unique()

df_sig
```

### Correlation between LPS at month 6 and cytokine at baseline
```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL")
LPS_data_M6_BL <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_M6_BL <- cytokine_data_BL[match(LPS_data_M6_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
n_M6BL_TB <- nrow(cytokine_data_M6_BL)
data_1 <- LPS_data_M6_BL%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_M6_BL%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_M6BL_TB <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p = df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_M6BL <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_M6BL <- df_sig$Cytokine%>%unique()

df_sig
```


## Calculate differential correlations 
### BLM6
```{r}
diff_BLBL <- diff_correlation_test(
  corr1=corr_BLBL_TB,
  corr2 = corr_M6M6_TB,
  n1 = n_BLBL_TB,
  n2 = n_M6M6_TB
)

diff_BLBL %>%
  filter(FDR < 0.05)
```

## TB-euglycemia {.tabset}
### Correlation between LPS at baseline and cytokines at baseline 

```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL", group=="Euglycemia")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL",newGroup=="Euglycemia")
LPS_data_BL <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_BL <- cytokine_data_BL[match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
n_BLBL_Euglycemia <- nrow(cytokine_data_BL)
data_1 <- LPS_data_BL%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_BL%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_BLBL_Euglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_BLBL_Euglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)
selected_cytokines_BLBL_Euglycemia <- df_sig$Cytokine%>%unique()
df_sig
```


### Correlation between LPS at month 6 and cytokines at month 6
```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6",group=="Euglycemia")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6",newGroup=="Euglycemia")
LPS_data_M6 <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_M6 <- cytokine_data_M6[match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
n_M6M6_Euglycemia <- nrow(cytokine_data_M6)
data_1 <- LPS_data_M6%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_M6%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_M6M6_Euglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_M6M6_Euglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_M6M6_Euglycemia <- df_sig$Cytokine%>%unique()
df_sig
```


### Correlation between LPS at baseline and cytokine at month 6
```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL",group=="Euglycemia")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6",newGroup=="Euglycemia")
LPS_data_BL_M6 <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_BL_M6 <- cytokine_data_M6[match(LPS_data_BL_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
n_BLM6_Euglycemia <- nrow(cytokine_data_BL_M6)
data_1 <- LPS_data_BL_M6%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_BL_M6%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_BLM6_Euglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_BLM6_Euglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_BLM6_Euglycemia <- df_sig$Cytokine%>%unique()

df_sig
```

### Correlation between LPS at month 6 and cytokine at baseline
```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6",group=="Euglycemia")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL",newGroup=="Euglycemia")
LPS_data_M6_BL <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_M6_BL <- cytokine_data_BL[match(LPS_data_M6_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
n_M6BL_Euglycemia <- nrow(cytokine_data_M6_BL)
data_1 <- LPS_data_M6_BL%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_M6_BL%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_M6BL_Euglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_M6BL_Euglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_M6BL_Euglycemia <- df_sig$Cytokine%>%unique()

df_sig
```



## TB-stress hyperglycemia {.tabset}
### Correlation between LPS at baseline and cytokines at baseline 

```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL", group=="Stress Hyperglycemia")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL",newGroup=="Stress Hyperglycemia")
LPS_data_BL <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_BL <- cytokine_data_BL[match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
n_BLBL_StressHyperglycemia <- nrow(cytokine_data_BL)
data_1 <- LPS_data_BL%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_BL%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_BLBL_StressHyperglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_BLBL_StressHyperglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)
selected_cytokines_BLBL_StressHyperglycemia <- df_sig$Cytokine%>%unique()
df_sig
```


### Correlation between LPS at month 6 and cytokines at month 6
```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6",group=="Stress Hyperglycemia")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6",newGroup=="Stress Hyperglycemia")
LPS_data_M6 <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_M6 <- cytokine_data_M6[match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
n_M6M6_StressHyperglycemia <- nrow(cytokine_data_M6)
data_1 <- LPS_data_M6%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_M6%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_M6M6_StressHyperglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_M6M6_StressHyperglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_M6M6_StressHyperglycemia <- df_sig$Cytokine%>%unique()
df_sig
```


### Correlation between LPS at baseline and cytokine at month 6
```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL",group=="Stress Hyperglycemia")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6",newGroup=="Stress Hyperglycemia")
LPS_data_BL_M6 <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_BL_M6 <- cytokine_data_M6[match(LPS_data_BL_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
n_BLM6_StressHyperglycemia <- nrow(cytokine_data_BL_M6)
data_1 <- LPS_data_BL_M6%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_BL_M6%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_BLM6_StressHyperglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_BLM6_StressHyperglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_BLM6_StressHyperglycemia <- df_sig$Cytokine%>%unique()

df_sig
```

### Correlation between LPS at month 6 and cytokine at baseline
```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6",group=="Stress Hyperglycemia")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL",newGroup=="Stress Hyperglycemia")
LPS_data_M6_BL <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_M6_BL <- cytokine_data_BL[match(LPS_data_M6_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
n_M6BL_StressHyperglycemia <- nrow(cytokine_data_M6_BL)
data_1 <- LPS_data_M6_BL%>%select(LPS)
data_1$LPS <- log10(data_1$LPS)
data_2 <- cytokine_data_M6_BL%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))

res <- rcorr(as.matrix(data_1), as.matrix(data_2), type = "spearman")

cor_mat <- res$r[
  colnames(data_1),
  colnames(data_2)
]

p_mat <- res$P[
  colnames(data_1),
  colnames(data_2)
]

df_long <- melt(cor_mat, varnames = c("Marker", "Cytokine"), value.name = "Correlation")
df_long$p.value <- melt(p_mat)$value
df_long$adj.p.value <- p.adjust(df_long$p.value,method = "fdr")
df_long$Cytokine <- row.names(df_long)
df_long$Marker <- "LPS"

corr_M6BL_StressHyperglycemia <- data.frame(from = df_long$Marker,
                    to   = df_long$Cytokine,
                    value = df_long$Correlation,
                    p=df_long$p.value)

df_sig <- df_long%>%filter(p.value<0.05)

links_M6BL_StressHyperglycemia <- data.frame(from = df_sig$Marker,
                    to   = df_sig$Cytokine,
                    value = df_sig$Correlation)

selected_cytokines_M6BL_StressHyperglycemia <- df_sig$Cytokine%>%unique()

df_sig
```




## Calculate differential correlations {.tabset}
### BLBL
```{r}
diff_BLBL <- diff_correlation_test(
  corr1=corr_BLBL_StressHyperglycemia,
  corr2 = corr_BLBL_Euglycemia,
  n1 = n_BLBL_StressHyperglycemia,
  n2 = n_BLBL_Euglycemia
)
```

### BLM6
```{r}
diff_BLM6 <- diff_correlation_test(
  corr1=corr_BLM6_StressHyperglycemia,
  corr2 = corr_BLM6_Euglycemia,
  n1 = n_BLM6_StressHyperglycemia,
  n2 = n_BLM6_Euglycemia
)
```

### M6BL
```{r}
diff_M6BL <- diff_correlation_test(
  corr1=corr_M6BL_StressHyperglycemia,
  corr2 = corr_M6BL_Euglycemia,
  n1 = n_M6BL_StressHyperglycemia,
  n2 = n_M6BL_Euglycemia
)
```

### M6M6
```{r}
diff_M6M6 <- diff_correlation_test(
  corr1=corr_M6M6_StressHyperglycemia,
  corr2 = corr_M6M6_Euglycemia,
  n1 = n_M6M6_StressHyperglycemia,
  n2 = n_M6M6_Euglycemia
)

```

## Make Figure 4:Chord plot for all cytokines in all participants

```{r}
df <- rbind(
  diff_BLBL %>% mutate(time = "baseline", time2 = "baseline"),
  diff_BLM6 %>% mutate(time = "baseline", time2 = "month 6"),
  diff_M6BL %>% mutate(time = "month 6", time2 = "baseline"),
  diff_M6M6 %>% mutate(time = "month 6", time2 = "month 6")
) %>%
  mutate(
    diff = r1 - r2,  
    edge_width = abs(diff),
    group = ifelse(diff > 0,
                   "stress-hyperglycemia",
                   "euglycemia")
  ) %>%
  filter(p_value < 0.05,(p1 < 0.05 | p2 < 0.05))


df$from_id <- paste0(df$from, "_", df$time)
df$to_id   <- paste0(df$to, "_", df$time2)

label_map <- c(
  setNames(df$from, df$from_id),
  setNames(df$to, df$to_id)
)

marker_BL <- sort(unique(df$from_id[grepl("_baseline$", df$from_id)]))
marker_M6 <- sort(unique(df$from_id[grepl("_month 6$", df$from_id)]))

cyto_BL <- sort(unique(df$to_id[grepl("_baseline$", df$to_id)]))
cyto_M6 <- sort(unique(df$to_id[grepl("_month 6$", df$to_id)]))

sector_order <- c(marker_BL, marker_M6, cyto_BL, cyto_M6)

grid.col <- c(
  setNames(rep("orange", length(marker_BL)), marker_BL),
  setNames(rep("lightcoral", length(marker_M6)), marker_M6),
  setNames(rep("yellowgreen", length(cyto_BL)), cyto_BL),
  setNames(rep("lightblue", length(cyto_M6)), cyto_M6)
)


group_cols <- c(
  "stress-hyperglycemia" = "#3abbcc",
  "euglycemia" = "#ca6332"
)

link_col <- group_cols[df$group]

link_lwd <- scales::rescale(df$edge_width, to = c(1, 8))

lgd_cor <- Legend(
  title = "Differential correlation strength (p-value < 0.05)",
  legend_gp = gpar(fill = c(
    group_cols["stress-hyperglycemia"],
    group_cols["euglycemia"]
  )),
  labels = c(
    "Stronger in TB-stress hyperglycemia",
    "Stronger in TB-euglycemia"
  ),
  title_gp = gpar(fontsize = 26, fontface = "bold"),
  labels_gp = gpar(fontsize = 26, fontface = "bold")
)

lgd_sectors <- Legend(
  title = "",
  at = c(
    "LPS markers (baseline)",
    "LPS markers (month 6)",
    "Cytokines (baseline)",
    "Cytokines (month 6)"
  ),
  legend_gp = gpar(fill = c(
    "orange",
    "lightcoral",
    "yellowgreen",
    "lightblue"
  )),
  labels_gp = gpar(fontsize = 26, fontface = "bold")
)

pdf("../Results/New Manuscript/Figure 4.pdf",
    width = 25, height = 30)

circos.clear()
circos.par(
  gap.degree = 2,
  start.degree = 90
)

par(mar = c(0,0,0,0))


chordDiagram(
  df[, c("from_id", "to_id", "diff")],
  order = sector_order,
  grid.col = grid.col,
  col = link_col,
  transparency = 0,
  link.lwd = link_lwd,
  annotationTrack = "grid",
  preAllocateTracks = 1
)


circos.trackPlotRegion(
  track.index = 1,
  panel.fun = function(x, y) {
    sector.name <- get.cell.meta.data("sector.index")
    
    circos.text(
      x = get.cell.meta.data("xcenter"),
      y = get.cell.meta.data("ylim")[1],
      labels = label_map[sector.name],
      facing = "clockwise",
      niceFacing = TRUE,
      adj = c(0, 0.5),
      cex = 3.5,
      font = 2
    )
  },
  bg.border = NA
)


draw(
  packLegend(lgd_sectors, lgd_cor),
  x = unit(0.85, "npc"),
  y = unit(0.85, "npc"),
  just = c("right", "top")
)

circos.clear()
dev.off()
```


## Scatter plots {.tabset}
### Scatterplot for significant associations between LPS at BL and cytokines at BL in TB only

```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL",Cohort=="MIDST")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL",newGroup!="Control")
LPS_data_BL <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_BL <- cytokine_data_BL[match(LPS_data_BL$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
LPS_data_BL$LPS <- log10(LPS_data_BL$LPS)
selected_cytokines_BLBL <- df%>%filter(time=="baseline",time2=="baseline")%>%pull(to)
links <- data.frame(from="LPS",to=selected_cytokines_BLBL)
groups <- LPS_data_BL$group

plot_data_BLBL <- pmap_dfr(
  links,
  function(from, to, ...) {
    
    if (from %in% "LPS" &&
        to %in% selected_cytokines_BLBL) {
      
      tibble(
        Marker = LPS_data_BL[[from]],
        Cytokine   = cytokine_data_BL[[to]],
        Group    = groups,
        Pair     = to
      )
    }
  }
)

plotBLBL <- plot_cyto_marker(plot_data_BLBL,method="spearman", 0)
pdf("../Results/New Manuscript/Figure 4-BLBL-spearman.pdf",width = 18,height = 6)
plotBLBL
dev.off()
```

### Scatterplot for significant associations between LPS at M6 and cytokines at M6

```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6")
LPS_data_M6 <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_M6 <- cytokine_data_M6[match(LPS_data_M6$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
LPS_data_M6$LPS <- log10(LPS_data_M6$LPS)

selected_cytokines_M6M6 <- df%>%filter(time=="month 6",time2=="month 6")%>%pull(to)
links <- data.frame(from="LPS",to=selected_cytokines_M6M6)
groups <- LPS_data_M6$group

plot_data_M6M6 <- pmap_dfr(
  links,
  function(from, to, ...) {
    
    if (from %in% "LPS" &&
        to %in% selected_cytokines_M6M6) {
      
      tibble(
        Marker = LPS_data_M6[[from]],
        Cytokine   = cytokine_data_M6[[to]],
        Group    = groups,
        Pair     = to
      )
    }
  }
)

plotM6M6 <- plot_cyto_marker(plot_data_M6M6,method="spearman", 0,n=6)
pdf("../Results/New Manuscript/Figure 4-M6M6-spearman.pdf",width = 18,height = 6)
plotM6M6
dev.off()
```

### Scatterplot for significant associations between LPS at BL and cytokines at M6

```{r}
LPS_data_BL <- LPS_data%>%filter(TIME=="BL")
cytokine_data_M6 <- cytokine_data%>%filter(TIME=="M6")
LPS_data_BL <- LPS_data_BL[which(!is.na(match(LPS_data_BL$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID))),]
cytokine_data_M6 <- cytokine_data_M6[match(LPS_data_BL$SAMPLE.ID,cytokine_data_M6$SAMPLE.ID),]
LPS_data_BL$LPS <- log10(LPS_data_BL$LPS)

selected_cytokines_BLM6 <- df%>%filter(time=="baseline",time2=="month 6")%>%pull(to)
links <- data.frame(from="LPS",to=selected_cytokines_BLM6)
groups <- LPS_data_BL$group

plot_data_BLM6 <- pmap_dfr(
  links,
  function(from, to, ...) {
    
    if (from %in% "LPS" &&
        to %in% selected_cytokines_BLM6) {
      
      tibble(
        Marker = LPS_data_BL[[from]],
        Cytokine   = cytokine_data_M6[[to]],
        Group    = groups,
        Pair     = to
      )
    }
  }
)

plotBLM6 <- plot_cyto_marker(plot_data_BLM6,method="spearman",0)
pdf("../Results/New Manuscript/Figure 4-BLM6-spearman.pdf",width = 18,height = 6)
plotBLM6
dev.off()
```


### Scatterplot for significant associations between LPS at M6 and cytokines at BL

```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6")
cytokine_data_BL <- cytokine_data%>%filter(TIME=="BL")
LPS_data_M6 <- LPS_data_M6[which(!is.na(match(LPS_data_M6$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID))),]
cytokine_data_BL <- cytokine_data_BL[match(LPS_data_M6$SAMPLE.ID,cytokine_data_BL$SAMPLE.ID),]
LPS_data_M6$LPS <- log10(LPS_data_M6$LPS)

selected_cytokines_M6BL <- df%>%filter(time=="month 6",time2=="baseline")%>%pull(to)
links <- data.frame(from="LPS",to=selected_cytokines_M6BL)
groups <- LPS_data_M6$group

plot_data_M6BL <- pmap_dfr(
  links,
  function(from, to, ...) {
    
    if (from %in% "LPS" &&
        to %in% selected_cytokines_M6BL) {
      
      tibble(
        Marker = LPS_data_M6[[from]],
        Cytokine   = cytokine_data_BL[[to]],
        Group    = groups,
        Pair     = to
      )
    }
  }
)

plotM6BL <- plot_cyto_marker(plot_data_M6BL,method="spearman", 0)
pdf("../Results/New Manuscript/Figure 4-M6BL-spearman.pdf",width = 18,height = 6)
plotM6BL
dev.off()
```

### Make Supplementary Figure 2 
```{r}
plot_data_BLBL <- plot_data_BLBL%>%mutate(time="baseline",time2="baseline")
plot_data_BLM6 <- plot_data_BLM6%>%mutate(time="baseline",time2="month 6")
plot_data_M6BL <- plot_data_M6BL%>%mutate(time="month 6",time2="baseline")
plot_data_M6M6 <- plot_data_M6M6%>%mutate(time="month 6",time2="month 6")
plot_data <- plot_data_BLBL%>%
  bind_rows(plot_data_BLM6)%>%
  bind_rows(plot_data_M6BL)%>%
  bind_rows(plot_data_M6M6)
plot <- plot_cyto_marker_with_time(plot_data,method="spearman",n=5,z=0)

pdf("../Results/New Manuscript/Supplementary Figure 2.pdf",width = 20,height = 18)
plot
dev.off()
```



### Make Supplementary Figure 3: Heatmap for LPS at baseline and cytokines at baseline
```{r}
LPS_data_BL <-LPS_data%>%filter(TIME=="BL", group!="Pre-Diabetes")
cytokine_data_BL <- cytokine_data_BL%>%filter(SAMPLE.ID%in%LPS_data_BL$SAMPLE.ID)
row.names(cytokine_data_BL) <- cytokine_data_BL$SAMPLE.ID
clinical_data_BL <- LPS_data_BL%>%select(SAMPLE.ID,group,LPS)
cytokine_data_BL <- cytokine_data_BL%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))%>%t()

pdf("../Results/New Manuscript/Supplementary Figure 3.pdf",width = 20,height = 15)
heatmapPlot_func(
  ClinicalData = clinical_data_BL,
  CytokineData = cytokine_data_BL,
  lps_col = "LPS"
)
dev.off()
```

### Make Supplementary Figure 4: Heatmap for LPS at month 6 and cytokines at month 6

```{r}
LPS_data_M6 <- LPS_data%>%filter(TIME=="M6", group!="Pre-Diabetes")
cytokine_data_M6 <- cytokine_data_M6%>%filter(SAMPLE.ID%in%LPS_data_M6$SAMPLE.ID)
row.names(cytokine_data_M6) <- cytokine_data_M6$SAMPLE.ID
clinical_data_M6 <- LPS_data_M6%>%select(SAMPLE.ID,group,LPS)
cytokine_data_M6 <- cytokine_data_M6%>%select(-c(SAMPLE.ID,TIME,newGroup,newGroup2))%>%t()
pdf("../Results/New Manuscript/Supplementary Figure 4.pdf",width = 20,height = 15)
heatmapPlot_func(
  ClinicalData = clinical_data_M6,
  CytokineData = cytokine_data_M6,
  lps_col = "LPS"
)
dev.off()
```

